---
title: "Notes_Lara"
output: html_document
date: "2022-08-04"
---
# general infos and tips
https://www.data-to-viz.com
https://datavizproject.com/
https://visualizationuniverse.com/charts/

## Data import
shows bike rentals 
```{r preparation, warning=FALSE, message=FALSE}
library(tidyverse)

bikes <- read_csv(
  "https://raw.githubusercontent.com/z3tt/graphic-design-ggplot2/main/data/london-bikes-custom.csv", 
  col_types = "Dcfffilllddddc"
)

bikes$season <- fct_inorder(bikes$season)

bikes
```

# aesthetics 

- we set constant propeties outside aes()
- … and map data-related properties inside aes()
- local settings and mappings override global properties
- grouping allows applying layers for subsets

shape is fixed and can be looked up online.
```{r}
ggplot(bikes, aes(x=temp_feel, y=temp))+
  geom_point(aes(fill=weather_type == "clear",
                 size = count),
                 shape = 23,   # makes a triangle 
                 color="black",
                 alpha = .5)

```

global aestethics: grouping indicates that geom smoothe is over the gouping variable day/night. 
```{r}
ggplot(
    bikes,
    aes(x = temp_feel, y = count,
        color = season,
        group = day_night)
  ) +
  geom_point(
    alpha = .5
  ) +
  geom_smooth(
    method = "lm",
    color = "black"
  )+
  geom_rug(
    alpha = .2
  )+
  labs(
    x = "Feels-like temperature (°F)",
    y = "Reported bike shares",
    title = "TfL bike sharing trends",
    color = "Season:",                                                           # titel of color legend
    subtitle = "Reported bike rents versus feels-like temperature in London",
    caption = "Data: TfL",
    tag = "Fig. 1"
  )+ theme_light(
  base_size = 14,
  base_family = "Courier-Italic" # schriftart
)+theme(
  panel.grid.minor = element_blank(),           # entfernt kleine Zwischenlinien
  plot.title = element_text(face = "bold"),    #macht titel fett
  plot.title.position = "plot",                # schiebt titel nach links 
  legend.position = "top")  

```
# statistical 

```{r}
data_summary <- function(x) {
   m <- mean(x)
   ymin <- m-sd(x)
   ymax <- m+sd(x)
   return(c(y=m,ymin=ymin,ymax=ymax))  # function to draw mean line 
}

ggplot(bikes, aes(x=reorder(weather_type, count), y = count))+   #reorders weather type by mean of count variable
  geom_jitter(aes(color=day_night), alpha=0.3)+ 
  geom_violin(alpha=0.6)+
  stat_summary(fun.data=data_summary, mult=1, 
                 geom="pointrange", width=0.8)+
  facet_wrap(
    ~ is_workday + day_night ,     # kann facet_wraps mit mehreren Konditionen gleichzeitig machen
    scales = "free_y"               # passt y scala bei jedem quadrant individuell an
  )
```

other mean to sort bars, if variable is a factor
```{r}
aes(y = fct_rev(        # from large to little
      fct_infreq(weather_type)
    ))

aes(y = fct_infreq(weather_type)) # from little to large
```

# theme

- style labels such as title, axis and legend texts with theme()
- format data-related labels with the labels argument of scale_*() y and x continous

```{r}
ggplot(
    bikes,
    aes(x = temp_feel, y = count,
        color = season)
  ) + 
  geom_point()+
  theme(
  plot.title = element_text(face = "bold"),  # change titel format
  plot.title.position = "plot",
  axis.text = element_text(   # change axis format
    color = "grey40",
    family = "Tabular",     
    face = "italic",
    colour = NULL,
    size = NULL,
    hjust = 1,
    vjust = 0,
    angle = 45,
    lineheight = 1.3, ## no effect here
    margin = margin(10, 0, 20, 0), ## no effect here
    debug = TRUE
  ),
  plot.tag = element_text(
    margin = margin(0, 12, -8, 0), ## trbl
    debug = TRUE   # fügt Farbe am Rand hinzu
  )
)+
  scale_y_continuous(
    breaks = 0:4*15000,
    labels = scales::comma_format(    # includes the word Bike into the y variable name and includes comma for 1000 indication
      suffix = " bikes"    
    ),
    name = NULL
  )+
  scale_x_discrete(
    name = NULL,
    labels = stringr::str_to_title # makes the labels to titel and delets variable name
  )




```
## chose own labels

- adjust strip text with the facet_*(labeller) functionality
```{r}
codes <- c(
  `TRUE` = "Workday",
  `FALSE` = "Weekend or Holiday"
)

 ggplot(
    bikes,
    aes(x = temp_feel, y = count,
        color = season)
  ) + 
  geom_point()+
  facet_wrap(
    ~ is_workday + day_night,
    labeller = labeller(
      day_night = stringr::str_to_title,           # change the titles so it is not weekday = weekend or holiday 
      is_workday = codes
    )
  )


```
## handling lond labels
```{r}
ggplot(
    bikes,
    aes(x = stringr::str_wrap(weather_type, 6),
        y = count)
  ) +
  geom_boxplot()

# or title

ggplot(
    bikes,
    aes(x = stringr::str_wrap(weather_type, 6),
        y = count)
  ) +
  geom_boxplot()+
  ggtitle("TfL bike sharing trends in 2015 and 2016 by season for day and night periods") +
  theme(
    plot.title =
      ggtext::element_textbox_simple(size = 20),
    plot.title.position = "plot"
  )


```
## plot titel
```{r}
g +
  ggtitle("TfL bike sharing trends in 2015 and 2016 by season for day and night periods") +
  theme(
    plot.title = ggtext::element_textbox_simple(  # fügt boxelement um den Titel hinzu
      margin = margin(t = 12, b = 12),
      padding = margin(rep(12, 4)),   # um die graue boy
      fill = "grey90",
      box.color = "grey40",
      r = unit(9, "pt"),
      halign = .5,
      face = "bold",
      lineheight = .9
    ),
    plot.title.position = "plot"
  )
```

# annotations

adding text bits into the plot

- add data-related annotations with geom_text|label()
- and data-unrelated annotations with annotate()

```{r}
ggplot(bikes, aes(humidity, temp)) +
  geom_point(size = 2, color = "grey") +
  annotate(
    geom = "text",
    x = c(90, 50),  # fügt zwei textbox hinzu
    y = c(27.5, 3.5),
    label = c("Text A", "Text B"),
    color = c("black", "firebrick"),
    size = c(5, 10),
    fontface = c("plain", "bold")
  )+
  annotate(            ## add a line
    geom = "segment",
    x = 90, xend = 82,
    y = 25, yend = 18.5
  )+ 
  annotate(               #adds a curved arrow
    geom = "curve",
    x = 90, xend = 82,
    y = 25, yend = 18.5,
    curvature = -.3,      # indicates how bent 
    angle = 140,          # angle of curve
    arrow = arrow())
```
annotate extreme points
```{r}
ggplot(
    filter(bikes, temp >= 27),    # will plot all points over 27 = outliers
    aes(x = humidity, y = temp,
        color = season == "summer")
  ) +
  geom_point(               
    data = bikes,                    # will plot the rest of the points
    color = "grey65", alpha = .3
  ) +
  geom_point(size = 2.5) +          # will plot outliers larger
  ggrepel::geom_text_repel(             #
    aes(label = str_to_title(season)),  # makes the labels to titels
    xlim = c(NA, 35), hjust = 1           ## force to the right
  ) +
  scale_color_manual(
    values = c("firebrick", "black"),
    guide = "none"
  ) +
  xlim(25, NA)
```
```{r}
ggplot(
    filter(bikes, temp > 20 & season != "summer"),  # filters for all temp higher 20 which are not summer
    aes(x = humidity, y = temp,
        color = season)
  ) +
  geom_point(
    data = bikes,                     #includes all the restpoints in grey
    color = "grey65", alpha = .3
  ) +
  geom_point() +                     # adds the outlier points
  ggforce::geom_mark_rect(
    aes(label = str_to_title(season))   #draws a rectangle around the outliers which are grouped by season
  ) +
  scale_color_brewer(
    palette = "Dark2",
    guide = "none"
  )
```

# add image into grid
```{r}
ggplot(bikes, aes(date, temp_feel)) +
  annotation_custom(
    grid::rasterGrob(
      image = img
    )
  ) +
  geom_point(color = "#71a5d4")
```


# scales

scale_objekt: kann x, y, farbe, form sein
scale_objekt_typ: kontinuierlich, diskret, zahl



    continuous(), discrete(), reverse(), log10(), sqrt(), date() for positions

    continuous(), discrete(), manual(), gradient(), gradient2(), brewer() for colors

    continuous(), discrete(), manual(), ordinal(), area(), date() for sizes

    continuous(), discrete(), manual(), ordinal() for shapes

    continuous(), discrete(), manual(), ordinal(), date() for transparency

careful when adjusting axis limits 
    scale_*_continuous(limits) removes the data outside the range
    coord_*(*lim) keeps the data and zooms into the range

```{r}
ggplot(
    bikes,
    aes(x = date, y = count,
        color = season)
  ) +
  geom_point() +
  scale_x_date(
    date_breaks = "4 months",
    date_labels = "%b '%y"
  ) +
  #scale_y_log10()  #dies würde die y achse logarithmieren 
  #scale_y_continuous(   #logarithmiert ebenfalls 
   # trans = "log10"
  #)+ 
  scale_y_continuous(
    breaks = c(0, seq(5000, 30000, by = 2500), 40000, 50000),  #macht zwischen 5000-30000 feinere Abstände
    limits = c(NA, 60000),  #Achsenlänge
    #expand = c(0, 0),  #wo ist das Kreuz
  ) +
  scale_color_discrete(
    name = "Season:",
    type = RColorBrewer::brewer.pal(
      n = 4, name = "Set1")  #überschreibt das color aes argument zu Beginn
  )

RColorBrewer::display.brewer.pal(n = 4, name = "Set1")

```
# circular coordinate system
```{r}
ggplot(
    filter(bikes, !is.na(weather_type)),
    aes(x = weather_type,
        fill = weather_type)
  ) +
  geom_bar() +
  coord_polar()

ggplot(
    filter(bikes, !is.na(weather_type)),
    aes(x = fct_infreq(weather_type),
        fill = weather_type)
  ) +
  geom_bar() +
  coord_polar(theta = "y")
```
## klassisches Kuchen Diagramm
```{r}
ggplot(
    filter(bikes, !is.na(weather_type)),
    aes(x = 1,
        fill = fct_rev(fct_infreq(weather_type)))
  ) +
  geom_bar(position = "stack") +
  coord_polar(theta = "y")
```

## klassisches geschichtetes Diagramm 
```{r}
ggplot(
    filter(bikes, !is.na(weather_type)),
    aes(x = 1,
        fill = fct_rev(fct_infreq(weather_type)))
  ) +
  geom_bar(position = "stack") +
  coord_cartesian()
```
## Lolipop Diagramm with two variables
```{r}
ggplot(bikes, aes(x = as.numeric(season), y = count, group = season)) +
  stat_summary(geom = "point", fun = "sum", size = 3) +                          # Summe der counts pro Season
  stat_summary(geom = "linerange", ymin = 0, fun.max = function(y) sum(y)) +     # Linie von 0 bis zur SUmme von Y pro Gruppe
  stat_summary(
    geom = "text", aes(label = season, y = 0),                                  # fügt Season Name bei Y=0 ein
    family = "Cabinet Grotesk", fontface = "bold", size = 4.5, hjust = 1.15
  ) +
  annotate(
    geom = "linerange", xmin = .7, xmax = 4.3, y = 0                            # fügt Linie auf Y=0 von x=0.7 bis 4.3 ein
  ) +
  coord_polar(theta = "y") +                                                    # dreht alles basierend auf der y Achse
  scale_x_discrete(expand = c(.5, .5)) +
  scale_y_continuous(limits = c(0, 7.5*10^6)) +                                 # vergrössert die y Achse, damit Punkte schön Abstand haben
  theme_void() +                                                               # Theme void nimmt alle Beschriftung raus
  theme(plot.margin = margin(rep(-130, 4)))
```
# legend position
```{r}
ggplot(
    bikes,
    aes(x = temp_feel, y = count,
        color = humidity)
  ) +
  geom_point() +
  scale_color_viridis_c(guide = guide_colorsteps(
      title.position = "top",       # puts the title of the legend on top
      title.hjust = .5,
      show.limits = TRUE,          # black frame around color interval in legend
      frame.colour = "black",
      frame.linewidth = 3,
      barwidth = unit(8, "lines")) +
  theme(
    legend.position = "bottom",
    legend.justification = "left" # shows were to go on the bottom
  )+ # or
  theme(
    legend.position = c(.25, .85),
    legend.direction = "horizontal"
  )


  )


```


# colors

predefined color sets
```{r}
RColorBrewer::display.brewer.all()
RColorBrewer::display.brewer.all(colorblindFriendly = TRUE)
rcartocolor::display_carto_all()
scico::scico_palette_show()
MetBrewer::display_all()
```
show difference via colorlighteness 

```{r}
ggplot(
    bikes, 
    aes(x = day_night, y = count)
  ) +
  geom_boxplot(
    aes(color = season,
        fill = after_scale(
          lighten(color, .8) # draws boxplot ligher
        ))
  ) +
  geom_jitter(
    aes(color = season,
        color = after_scale(
          darken(color, .3)   # draws points darker
        )), 
    position = position_jitterdodge(
      dodge.width = .75, 
      jitter.width = .2
    ),
    alpha = .4
  ) 
```
## create new palets 
```{r}
ggplot(
    bikes, 
    aes(x = temp_feel, y = count, 
        color = temp_feel)
  ) +
  geom_point() +
  scale_color_gradient(
    low = "#28A87D",
    high = "#FFD166"
  )


ggplot(
    bikes, 
    aes(x = temp_feel, y = count, 
        color = temp_feel)
  ) +
  geom_point() +
  scale_color_gradient2(
    low = "#663399",
    high = "#993334",
    mid = "grey92",
    midpoint = 10     # defines the middlepoint and what color it should have 
  )
```


# systemfonts 
register_variant(): This function is a wrapper around register_font() that allows you to easily create variants of existing system fonts, e.g. to target different weights and/or widths, or for attaching OpenType features to a font. 

system_fonts(): shows all system fonts

```{r}
systemfonts::system_fonts() %>%   
  filter(str_detect(family, "Courier")) %>%
  pull(name, weight) %>%
  sort()


tst <- systemfonts::register_variant(
  name = "Courier-Italic",
  family = "Courier",
  weight = "normal",
  features = systemfonts::font_feature(letters = "stylistic")
)
```

# patchwork

combine multiple ggplots

library(patchwork)

{patchwork} is the most modern and simplest way to compose plots

    use +, / and | to combine plots
    use plot_layout() to adjust the layout, widths and heights
    use plot_annotation() to add tags and general labels
    use inset_element() to add inset graphics


```{r}
library(patchwork)
(p1 + p2) / p3   # two on top, one on the bottom

(p1 + p2) / p3 & theme(legend.position = "none", plot.background = element_rect(color = "black", size = 3)) # draws black lines around plots
```

## overall titel
```{r}
(pl1 + pl2) / pl3 +
plot_annotation(tag_levels = "1", tag_prefix = "P", title = "An overarching title for all 3 plots, placed on the very top while all other titles are sitting below the tags.")
```

## add plot into plot
```{r}
pl1 + inset_element(pl2, l = .6, b = .1, r = 1, t = .6)
```

