---
title: "Notes_Lara"
output: html_document
date: "2022-08-04"
---
# general infos and tips
https://www.data-to-viz.com
https://datavizproject.com/
https://visualizationuniverse.com/charts/

## Data import
shows bike rentals 
```{r preparation, warning=FALSE, message=FALSE}
library(tidyverse)

bikes <- read_csv(
  "https://raw.githubusercontent.com/z3tt/graphic-design-ggplot2/main/data/london-bikes-custom.csv", 
  col_types = "Dcfffilllddddc"
)

bikes$season <- fct_inorder(bikes$season)

bikes
```

# aesthetics 

- we set constant propeties outside aes()
- … and map data-related properties inside aes()
- local settings and mappings override global properties
- grouping allows applying layers for subsets

shape is fixed and can be looked up online.
```{r}
ggplot(bikes, aes(x=temp_feel, y=temp))+
  geom_point(aes(fill=weather_type == "clear",
                 size = count),
                 shape = 23,   # makes a triangle 
                 color="black",
                 alpha = .5)

```

global aestethics: grouping indicates that geom smoothe is over the gouping variable day/night. 
```{r}
ggplot(
    bikes,
    aes(x = temp_feel, y = count,
        color = season,
        group = day_night)
  ) +
  geom_point(
    alpha = .5
  ) +
  geom_smooth(
    method = "lm",
    color = "black"
  )+
  geom_rug(
    alpha = .2
  )+
  labs(
    x = "Feels-like temperature (°F)",
    y = "Reported bike shares",
    title = "TfL bike sharing trends",
    color = "Season:",                                                           # titel of color legend
    subtitle = "Reported bike rents versus feels-like temperature in London",
    caption = "Data: TfL",
    tag = "Fig. 1"
  )+ theme_light(
  base_size = 14,
  base_family = "Courier-Italic" # schriftart
)+theme(
  panel.grid.minor = element_blank(),           # entfernt kleine Zwischenlinien
  plot.title = element_text(face = "bold"),    #macht titel fett
  plot.title.position = "plot",                # schiebt titel nach links 
  legend.position = "top")  

```
# statistical 

```{r}
data_summary <- function(x) {
   m <- mean(x)
   ymin <- m-sd(x)
   ymax <- m+sd(x)
   return(c(y=m,ymin=ymin,ymax=ymax))  # function to draw mean line 
}

ggplot(bikes, aes(x=reorder(weather_type, count), y = count))+   #reorders weather type by mean of count variable
  geom_jitter(aes(color=day_night), alpha=0.3)+ 
  geom_violin(alpha=0.6)+
  stat_summary(fun.data=data_summary, mult=1, 
                 geom="pointrange", width=0.8)+
  facet_wrap(
    ~ is_workday + day_night ,     # kann facet_wraps mit mehreren Konditionen gleichzeitig machen
    scales = "free_y"               # passt y scala bei jedem quadrant individuell an
  )
```

other mean to sort bars, if variable is a factor
```{r}
aes(y = fct_rev(        # from large to little
      fct_infreq(weather_type)
    ))

aes(y = fct_infreq(weather_type)) # from little to large
```


# scales

scale_objekt: kann x, y, farbe, form sein
scale_objekt_typ: kontinuierlich, diskret, zahl



    continuous(), discrete(), reverse(), log10(), sqrt(), date() for positions

    continuous(), discrete(), manual(), gradient(), gradient2(), brewer() for colors

    continuous(), discrete(), manual(), ordinal(), area(), date() for sizes

    continuous(), discrete(), manual(), ordinal() for shapes

    continuous(), discrete(), manual(), ordinal(), date() for transparency

careful when adjusting axis limits 
    scale_*_continuous(limits) removes the data outside the range
    coord_*(*lim) keeps the data and zooms into the range

```{r}
ggplot(
    bikes,
    aes(x = date, y = count,
        color = season)
  ) +
  geom_point() +
  scale_x_date(
    date_breaks = "4 months",
    date_labels = "%b '%y"
  ) +
  #scale_y_log10()  #dies würde die y achse logarithmieren 
  #scale_y_continuous(   #logarithmiert ebenfalls 
   # trans = "log10"
  #)+ 
  scale_y_continuous(
    breaks = c(0, seq(5000, 30000, by = 2500), 40000, 50000),  #macht zwischen 5000-30000 feinere Abstände
    limits = c(NA, 60000),  #Achsenlänge
    #expand = c(0, 0),  #wo ist das Kreuz
  ) +
  scale_color_discrete(
    name = "Season:",
    type = RColorBrewer::brewer.pal(
      n = 4, name = "Set1")  #überschreibt das color aes argument zu Beginn
  )

RColorBrewer::display.brewer.pal(n = 4, name = "Set1")

```
# circular coordinate system
```{r}
ggplot(
    filter(bikes, !is.na(weather_type)),
    aes(x = weather_type,
        fill = weather_type)
  ) +
  geom_bar() +
  coord_polar()

ggplot(
    filter(bikes, !is.na(weather_type)),
    aes(x = fct_infreq(weather_type),
        fill = weather_type)
  ) +
  geom_bar() +
  coord_polar(theta = "y")
```
## klassisches Kuchen Diagramm
```{r}
ggplot(
    filter(bikes, !is.na(weather_type)),
    aes(x = 1,
        fill = fct_rev(fct_infreq(weather_type)))
  ) +
  geom_bar(position = "stack") +
  coord_polar(theta = "y")
```

## klassisches geschichtetes Diagramm 
```{r}
ggplot(
    filter(bikes, !is.na(weather_type)),
    aes(x = 1,
        fill = fct_rev(fct_infreq(weather_type)))
  ) +
  geom_bar(position = "stack") +
  coord_cartesian()
```
## Lolipop Diagramm with two variables
```{r}
ggplot(bikes, aes(x = as.numeric(season), y = count, group = season)) +
  stat_summary(geom = "point", fun = "sum", size = 3) +                          # Summe der counts pro Season
  stat_summary(geom = "linerange", ymin = 0, fun.max = function(y) sum(y)) +     # Linie von 0 bis zur SUmme von Y pro Gruppe
  stat_summary(
    geom = "text", aes(label = season, y = 0),                                  # fügt Season Name bei Y=0 ein
    family = "Cabinet Grotesk", fontface = "bold", size = 4.5, hjust = 1.15
  ) +
  annotate(
    geom = "linerange", xmin = .7, xmax = 4.3, y = 0                            # fügt Linie auf Y=0 von x=0.7 bis 4.3 ein
  ) +
  coord_polar(theta = "y") +                                                    # dreht alles basierend auf der y Achse
  scale_x_discrete(expand = c(.5, .5)) +
  scale_y_continuous(limits = c(0, 7.5*10^6)) +                                 # vergrössert die y Achse, damit Punkte schön Abstand haben
  theme_void() +                                                               # Theme void nimmt alle Beschriftung raus
  theme(plot.margin = margin(rep(-130, 4)))
```



# systemfonts 
register_variant(): This function is a wrapper around register_font() that allows you to easily create variants of existing system fonts, e.g. to target different weights and/or widths, or for attaching OpenType features to a font. 

system_fonts(): shows all system fonts

```{r}
systemfonts::system_fonts() %>%   
  filter(str_detect(family, "Courier")) %>%
  pull(name, weight) %>%
  sort()


tst <- systemfonts::register_variant(
  name = "Courier-Italic",
  family = "Courier",
  weight = "normal",
  features = systemfonts::font_feature(letters = "stylistic")
)
```

